---
title: "Wk5Prac"
output: html_document
date: "2022-11-02"
---

GOAL: join polygons of boroughs with number of hotels and airbnbs and hotels

##ONE: Load Libraries and Data

```{r setup, include=FALSE}
library(sf)
library(tmap)
library(tmaptools)
library(tidyverse)
library(here)
library(leafpop)
library(leaflet)
library(grid)
library(readr)
library(janitor)
library(dplyr)
library(grid)
library(OpenStreetMap)

airbnb <- read_csv(here::here("wk5_data", "listings.csv")) %>%
  st_as_sf(., coords = c("longitude", "latitude"), crs = 4326) %>%
  st_transform(., 27700 )%>%
  filter(room_type == 'Entire home/apt' & availability_365 =='365')

Londonborough <- st_read(here::here("wk5_data","statistical-gis-boundaries-london", "ESRI", "London_Borough_Excluding_MHW.shp")) %>%
  st_transform(., 27700)

OSM <- st_read(here::here("wk5_data","greater-london-latest-free.shp", "gis_osm_pois_a_free_1.shp")) %>% 
  st_transform(.,27700) %>% 
  filter(fclass=='hotel')

Worldcities <- st_read(here::here("wk5_data", "World_Cities", "World_Cities.shp")) %>%
  st_transform(., 27700)

UK_outline <- st_read(here::here("wk5_data", "gadm41_GBR_shp", "gadm41_GBR_0.shp")) %>%
  st_transform(., 27700)

UkSimple <- UK_outline %>% 
  st_simplify(., dTolerance = 1000) %>%
  st_geometry()

UkSimple
plot(UkSimple)
```

## Wk5 Prac Overview
Join and Manipulate Data
1. Spatial Join using st_join()
2. Write general join function using functions we'll learn in week 7, replace data 1 and data 2 are what we want to join)
3. Join both hotels and airbnbs to london borough maps, that will give each hotel its own row so the borough row names will be repeated
4. Filter

```{r}

join_example <- st_join(Londonborough, OSM)

joinfun <-  function(data1, data2){
  output <- data1 %>% 
    st_join(data2,.) %>%
    add_count(GSS_CODE, name = "hotel_in_borough")
  return(output)}

Hotels <- joinfun(OSM, Londonborough)
Airbnbs <- joinfun(airbnb, Londonborough)
Worldcities2 <- Worldcities %>% 
  filter(CNTRY_NAME=='United Kingdom'& Worldcities$CITY_NAME=='Birmingham'| Worldcities$CITY_NAME=='London'| Worldcities$CITY_NAME=='Edinburgh')

#Here, we use group_by and summarise so each borough will have a row with the count of airbnbs and hotels

Hotels <- Hotels %>%
  group_by(., GSS_CODE, NAME)%>%
  summarise('Accomodation count' = unique(hotel_in_borough))

Airbnbs <- Airbnbs %>%
  group_by(., GSS_CODE, NAME)%>%
  summarise('Accomodation count' = unique(hotel_in_borough))

#Crop the UK Map
  newbb <- c(xmin=-296000, ymin=5408, xmax=655696, ymax=1000000)
#(I am guessing bb stands for bounding box)
  UK_outlinecrop <- UkSimple %>%
    st_crop(., newbb)
```

Here, we will use tmap to make a few types of maps: Inset map, Basic interactive map, Advanced interactive map

##Inset Map, Pt. 1 - this is the four quadrant map Andy discussed in the practical, each of the objects below

```{r}

tmap_mode("plot")

# set the breaks for our mapped data
breaks = c(0, 5, 12, 26, 57, 286) 

# make the object for each quandrant of the map
tm1 <- tm_shape(Hotels) + tm_polygons("Accomodation count", breaks=breaks,palette="PuBu") +  tm_legend(show=FALSE) + tm_layout(frame=FALSE) + tm_credits("(a)", position=c(0,0.85), size=1.5)

tm2 <- tm_shape(Airbnbs) + tm_polygons("Accomodation count", breaks=breaks,palette="PuBu") + tm_legend(show=FALSE) + tm_layout(frame=FALSE) + tm_credits("(b)", position=c(0,0.85), size=1.5)

tm3 <- tm_shape(UK_outlinecrop) + tm_polygons(col="darkslategray1") + tm_layout(frame=FALSE) + tm_shape(Worldcities2) + tm_symbols(col = "red", scale = .5) + tm_text("CITY_NAME", xmod=-1, ymod=-0.5)

legend <- tm_shape(Hotels) + tm_polygons("Accomodation count", palette="PuBu") + tm_scale_bar(position=c(0.2,0.04), text.size=0.6) + tm_compass(north=0, position=c(0.65,0.6)) + tm_layout(legend.only = TRUE, legend.position=c(0.2,0.25),asp=0.1) + tm_credits("(c) OpenStreetMap contrbutors and Air b n b", position=c(0.0,0.0))
  
t=tmap_arrange(tm1, tm2, tm3, legend, ncol=2)
#Q: can you do this with any number of "tm1" objects?
#A: Yes

#Here, you use the grid package to do the same thing

grid.newpage()
#Q: what does this do?
#A: it erases the current device, but you don't really need this, might be useful when exporting.

pushViewport(viewport(layout=grid.layout(2,2)))
print(tm1, vp=viewport(layout.pos.col=1, layout.pos.row=1, height=5))
print(tm2, vp=viewport(layout.pos.col=2, layout.pos.row=1, height=5))
print(tm3, vp=viewport(layout.pos.col=1, layout.pos.row=2, height=5))
print(legend, vp=viewport(layout.pos.col=2, layout.pos.row=2, height=5))
```


##Inset Map, Pt. 2
```{r}
Londonbb = st_bbox(Airbnbs, crs = st_crs(Airbnbs)) %>% 
  st_as_sfc()
#Q: what is this function/what is it doing?
#A: Take geometry and make it into an sfc object, a geometry collection, it makes it a polygon

main <- tm_shape(Airbnbs, bbbox = Londonbb) + 
  tm_polygons("Accomodation count", breaks=breaks, palette="PuBu")+
  tm_scale_bar(position = c("left", "bottom"), text.size = .75)+
  tm_layout(legend.position = c("right","top"), legend.text.size=.75, legend.title.size = 1.1, frame=FALSE)+
  tm_credits("(c) OpenStreetMap contrbutors and Airbnb", position=c(0.0,0.0))+
  #tm_text(text = "NAME", size = .5, along.lines =T, remove.overlap=T,  auto.placement=F)+
  tm_compass(type = "8star", position = c(0.06, 0.1)) +

  #bottom left top right
  tm_layout(inner.margin=c(0.02,0.02,0.02,0.2))

inset = tm_shape(UK_outlinecrop) + tm_polygons() +
  tm_shape(Londonbb)+ 
  tm_borders(col = "grey40", lwd = 3)+
  tm_layout(frame=FALSE, bg.color = "transparent")+
  tm_shape(Worldcities2) +
  tm_symbols(col = "red", scale = .5)+
  tm_text("CITY_NAME", xmod=-1.5, ymod=-0.5)

main
print(inset, vp = viewport(0.86, 0.29, width = 0.5, height = 0.55))

tmap_save(t, 'hotelsandairbnbR.png')

library(grid)
tmap_save(main,insets_tm = inset,insets_vp=viewport(x=0.86, y=0.29, width=.5, height=.55), filename="test.pdf", dpi=600)
```


##Basic Interactive Map
```{r}
tmap_mode("view")

tm_shape(Airbnbs) + 
  tm_polygons("Accomodation count", breaks=breaks) 
```


##Advanced Interactive Map

```{r}
#libraries for pop-up boxes: leaflet, leafpop

#join data
Joined <- Airbnbs %>% 
  st_join(., Hotels, join = st_equals) %>% 
  dplyr::select(GSS_CODE.x, NAME.x, 'Accomodation count.x', 'Accomodation count.y') %>% 
  dplyr::rename('GSS code' = 'GSS_CODE.x', 'Borough' = 'NAME.x', 'Airbnb count'= 'Accomodation count.x', 'Hotel count' = 'Accomodation count.y') %>% 
  st_transform(., 4326)
  
#remove the geometry from the pop-ups to avoid them impact the mapping

popupairbnb <- Joined %>% 
  st_drop_geometry() %>% 
  dplyr::select('Airbnb count', Borough) %>% 
  popupTable()

popuphotel <- Joined %>% 
  st_drop_geometry() %>% 
  dplyr::select('Hotel count', Borough) %>% 
  popupTable()

tmap_mode("view")

#set colour palette
pal1 <- Joined %>%
  colorBin(palette = "YlOrRd", domain=.$'Airbnb count', bins=breaks)

pal1 <-colorBin(palette = "YlOrRd", domain=Joined$'Airbnb count', bins=breaks)

pal2 <- Joined %>%
  colorBin(palette = "YlOrRd", domain=.$'Hotel count', bins=breaks)

#make the actual map
map<- leaflet(Joined) %>%
  # add basemap options
  addTiles(group = "OSM (default)") %>%
  addProviderTiles(providers$Stamen.Toner, group = "Toner") %>%
  addProviderTiles(providers$Stamen.TonerLite, group = "Toner Lite") %>%
  addProviderTiles(providers$CartoDB.Positron, group = "CartoDB") %>%
  # add polygons
  addPolygons(color="white", 
              weight = 2,
              opacity = 1,
              dashArray = "3",
              popup = popupairbnb,
              fillOpacity = 0.7,
              fillColor = ~pal2(`Airbnb count`),
              group = "Airbnbs")%>%
  
  addPolygons(fillColor = ~pal2(`Hotel count`), 
              weight = 2,
              opacity = 1,
              color = "white",
              dashArray = "3",
              popup = popupairbnb,
              fillOpacity = 0.7,group = "Hotels")%>% 
  
  #add legend
  addLegend(pal=pal2, values='Hotel count', group=c("Airbnbs", "Hotels"), position="bottomleft", title='Accomodation Count') %>% 
  # specify layers control
  addLayersControl(baseGroups = c("OSM (default)", "Toner", "Toner Lite", "CartoDB"), overlayGroups = c("Airbnbs", "Hotels"), options = layersControlOptions(collapsed = FALSE))

# plot the map
map
```


